---
title: "Create a Report for a Phenotype Dataset"
output:
  rmarkdown::html_vignette:
    highlight: pygments
    toc: true
    fig_width: 5
    fig.align: "center"
    df_print: !expr function(df) {print(df, row.names = FALSE, max = 20 * ncol(df))}
vignette: >
  %\VignetteIndexEntry{Create a Report for a Phenotype Dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  dev = "png"
)
```

```{r setup}
library(process.phenotypes)
```

## Overview

The primary intended use of this package is to generate a cleaned, structured output dataset
alongside an html report describing the contents of the input file and the results of
the processing conducted by the package. This functionality is contained in the function
`create.phenotype.report`. To use this function, the following things are required:

- an input dataset
  - this should be plaintext, tab- or comma-delimited. If it can be read by `read.table`,
    it should be fine. If the input dataset is jagged in some way (for example, comma-delimited
	input that also contains commas within individual fields, and the fields are not quoted),
	some preprocessing outside of the package may be required.
- a dataset configuration file
  - this file is discussed at some length in [the manual configuration vignette](manual-configuration.html),
    or for users with SurveyCTO datasets, in [the SurveyCTO conversion vignette](configuration-from-SurveyCTO-form.html).
- a shared models configuration file
  - this file is discussed both in the aforementioned vignettes and in [the dedicated vignette regarding shared models](shared-models.html).
- an output filename for the resulting html report, with the required suffix `.html`

## Usage

Extensive documentation and discussion of the accepted options and cleaning steps of `create.phenotype.report` are provided
in both [the readthedocs documentation](http://54gene-processphenotypes-docs.s3-website.us-east-2.amazonaws.com/index.html)
and the internal R help documentation `?create.phenotype.report`.

As a very brief example:

```{r acquire.datasets, echo=TRUE, eval=TRUE}
input.data <- system.file("examples", "example.data.tsv",
  package = "process.phenotypes"
)
dataset.config <- system.file("examples", "example.dataset.yaml",
  package = "process.phenotypes"
)
shared.models <- system.file("examples", "example.shared_models.yaml",
  package = "process.phenotypes"
)
output.html <- tempfile("create_phenotype_report", fileext = "html")
```

These files contain the following:

```{r show.input.data, eval=TRUE, echo=FALSE}
cat("## example.data.tsv\n\n")
read.table(input.data, sep = "\t", header = TRUE)
```

```{r, echo=FALSE}
# nolint start
```

```{yaml}
## example.dataset.yaml
tag: HW
globals:
  min_age_for_inclusion: 18
  max_invalid_datatypes_per_subject: 10
  consent_inclusion_file: ~
  consent_exclusion_file: ~
variables:
  HW0001:
    name: "subject.ids"
    canonical_name: "participant ID"
    type: "string"
    subject_id: yes
  HW0002:
    name: "age"
    canonical_name: "participant age"
    type: "numeric"
    subject_age: yes
  HW0003:
    name: "date of birth"
    canonical_name: "subject date of birth (year)"
    type: "date"
  HW0004:
    name: "height"
    canonical_name: "subject height (meters)"
    type: "numeric"
  HW0005:
    name: "sex"
    canonical_name: "subject self-reported sex"
    shared_model: "sex"
  HW0006:
    name: "blood pressure"
    canonical_name: "subject blood pressure"
    type: "bp"
  HW0007:
    name: "satisfaction rating"
    canonical_name: "how much does subject like a thing"
    shared_model: "satis_rating"
  HW0008:
    name: "weight"
    canonical_name: "subject weight (kg)"
    type: "numeric"
derived:
  bmi_derived:
    name: "bmi, computed from self-reported height and weight"
    type: "numeric"
    code: |
      HW0008 / HW0004 ^ 2
```

```{yaml}
## example.shared_models.yaml
models:
  sex:
    type: "categorical"
    levels:
      "1":
        name: "male"
        alternate_patterns:
          - "^[mM]$"
          - "^Male$"
      "2":
        name: "female"
        alternate_patterns:
          - "^[fF]$"
          - "^Female"
  satis_rating:
    type: "ordinal"
    levels:
      "1":
        name: "dissatisfied"
      "2":
        name: "neutral"
      "3":
        name: "satisfied"
```

```{r, echo=FALSE}
# nolint end
```

To generate your output report, and optionally the processed data in tsv or a variety of other
accepted formats, run the following:

```{r run.create.phenotype.report, echo=TRUE, eval=TRUE}
create.phenotype.report(input.data,
  dataset.config,
  shared.models,
  output.html,
  sep = "\t",
  quote = "",
  write.tsv = TRUE
)
```

## A General Workflow

The process of dataset cleaning can be rather complicated, and is usually very iterative.
Our usual workflow looks like this:

- generate the above configuration files with placeholder entries for most variables
  - this usually entails minimalist settings that preserve input data, as follows:
  
```{yaml}
variables:
  HW00001:
    name: "varname1"
    type: "string"
    canonical_name: "my variable description 1"
  HW00002:
    name: "varname2"
    type: "string"
    canonical_name: "my variable description 2"
```

and so on.

- optionally, validate the configuration data you've generated with the package's
  [built-in validation utility](yaml-validation.html).
- run `create.phenotype.report` with the simple configuration. This will generate basic data descriptions
  that can inform your configuration refinement.
  - the report will, by default, suppress tabular reporting information for variables with more unique
    values than `{sample size} / 3`. If that affects your variables, you can consider either inspecting
	the variables in R separately, or adjusting the cutoff factor with the corresponding parameter
	in `create.phenotype.report` (see manual in `?create.phenotype.report` for all available settings)
- update the configuration to reflect more refined settings.
- and repeat the above as needed.
