---
title: "derived-variables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{derived-variables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(process.phenotypes)
```

# Creation of Derived Variables

## Overview

Derived variables are any variables that were not included in the primary data
collection from the subject, but are generated using that data.  Examples can include:

- Splitting blood pressure from SBP/DBP to two new variables, one each for systolic and diastolic
- Calculating BMI from height and weight
- Performing unit conversion
- Normalizing data, e.g. with an inverse normal transform
- Creating binary encodings from free text entries, e.g. determining whether a patient has had
  treatment with hydroxyurea by searching for a string match in a field containing treatment notes
- Harmonizing variables for downstream merging across multiple datasets, e.g. creating a standardized
  name "subject_age" to map the age variable in all datasets regardless of the original variable name

## Example Use Cases

Let's start with a baseline YAML configuration and add derived variables for a variety of use cases.

```{yaml}

variables:
  VAR00001:
    name: "SubjID"
    canonical_name: "Subject ID"
    type: "string"
    subject_id: "yes"
  VAR00002:
    name: "BP"
    canonical_name: "blood pressure (SBP/DBP)"
    type: "bp"
  VAR00003:
    name: "What is your height?"
    canonical_name: "patient's height in cm"
    type: "numeric"
  VAR00004:
    name: "What is your weight?"
    canonical_name: "patient's weight in kg"
    type: "numeric"
  VAR00005:
    name: "treatment_notes"
    canonical_name: "patient treatment notes from medical records"
    type: "string"
  VAR00006:
    name: "patient age"
    canonical_name: "age in years"
    type: "numeric"

```

The derived variables, also mentioned [here](manual-configuration.html), are entered in their
own yaml block at the same level as the `variables` section.  Derived variables are created after
all the cleaning steps in `process.phenotypes` have been performed, and are subsequently also
cleaned according to their declared type.  You can derive variables from other derived variables;
the package will iterate until all variables have been accounted for.

### Deriving SBP and DBP

We can see that `VAR00002` is of type `bp`, which will result in values like `120/90` after
cleaning (see the manual configuration [vignette](manual-configuration.html) for more details).
You may want to separate these into two variables for downstream analysis.  This can be accomplished
by adding derived variables as follows:

```{yaml}

derived:
  VAR00002_SBP:
    name: "derived_SBP"
    type: "numeric"
    code: as.numeric(stringr::str_replace(VAR00002, "/.*", ""))
  VAR00002_DBP:
    name: "derived_DBP"
    type: "numeric"
    code: as.numeric(stringr::str_replace(VAR00002, ".*/", ""))
```

Note that the text in the `code` block is executed in an
isolated environment and does not affect the underlying original data.  You have access to
all of the variables in the dataset as vectors, labeled as their user-defined names (e.g.
"VAR00002").  The derived variables as written above will be emitted in the cleaned output
phenotype matrix.

### Unit Conversion

```{yaml}

derived:
  height_in_meters:
    name: "derived_height_in_meters"
    type: "numeric"
    code: VAR00003 / 100
```

### Calculate BMI

```{yaml}

derived:
  calculated_bmi:
    name: "derived_BMI"
    type: "numeric"
    code: VAR00004 / height_in_meters ^ 2
```

### Normalize Data

```{yaml}

derived:
  VAR00003_INT:
    name: "height, inverse normal transform"
    type: "numeric"
    code: derive.rank.normal.transform(VAR00003)

```

Note the use of a convenience function here, provided in `process.phenotypes`.

### Binary Encodings from Free Text

```{yaml}
derived:
  HU_use:
    name: "Treatment with hydroxyurea"
    shared_model: "yesno"
    code: factor(ifelse(stringr::str_detect(VAR00005, "hydroxyurea"), "yes", "no"))

```

### Harmonizing Variable Names

```{yaml}

derived:
  std_height:
    name: "standardized_height"
    type: "numeric"
    code: VAR00003
```
