---
title: "Using shared models to harmonize variables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using shared models to harmonize variables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(process.phenotypes)
```

# Shared Model Configuration

## Overview

Variables scanned by `process.phenotypes` can be configured very flexibly, with
a wide variety of configuration settings. However, there are pitfalls to certain
methods of configuration. Most notably, it can be difficult to keep similar
variables in the same dataset or even across datasets configured in the same,
or at least compatible, ways.

The package attempts to address this kind of desynchronization by the use of
what we call "shared models": configuration data that is specified once and then
referenced by a user-specified alias. These models are configured in a separate
YAML configuration file that is then provided to `create.phenotype.report`.
This vignette will discuss some of the approaches to shared model specification,
and how they might be leveraged to facilitate downstream analysis.

## General Structure of a Shared Model

Shared model configuration is very similar to standard variable configuration.
Consider the following example:

```{yaml}
models:
  yesno:
    type: "categorical"
    levels:
      "1":
        name: "no"
        alternate_patterns:
          - "0"
      "2":
        name: "yes"
        alternate_patterns:
          - "1"
    "na-values":
      - "none"
      - "unknown"
      - "not available"
```

Please note the following important characteristics of the above configuration:

- Every shared model is specified as a mapping under the top level `models` key.
- `models` should be specified exactly once in a given shared models file.
- The tag `yesno` is the name of one particular shared model. When using this shared
  model in a variable configuration (see below), this name is what is used to pull
  in this shared model. The names of models should be unique, but otherwise are
  flexibly user-specified.
- The configuration data under `yesno` corresponds to the usual variable configuration
  settings, with one important exception: there should not be a `name` key/value pair
  in any shared model. The name is meant to be inherited from each individual variable
  configuration that uses this shared model.


Shared models most naturally fit categorical variables, but are not necessarily required
to be categorical. Any accepted `process.phenotypes` variable configuration can be
used in a shared model:

```{yaml}
models:
  height:
    type: "numeric"
    bounds:
      min: 0.75
      max: 2.25
```

Aliasing the statement `type: "numeric"` under a shared model is a bit excessive,
but see that the `bounds` settings can also be inherited using shared models.
In this way, if you have multiple variables that contain, for example, human 
height in meters, you can set a single set of numeric limits on those variables,
and then if you need to adjust them, you can adjust them exactly once and the
change will be inherited by all relevant variables.

## Using a Shared Model in a Dataset Configuration

Once you have created some shared models, you can use them in a dataset
configuration file as follows:

```{yaml}
tag: "HW"
globals:
  min_age_for_inclusion: 18
  max_invalid_datatypes_per_subject: 10
  consent_inclusion_file: ~
  consent_exclusion_file: ~
variables:
  HW00001:
    name: "subject_id"
    type: "string"
    suppress_reporting: yes
    subject_id: yes
  HW00002:
    name: "subject_height"
    shared_model: "height"
  HW00003:
    name: "subject_likes_cake"
    shared_model: "yesno"
```

The above variables `HW00002` and `HW00003` are completely specified.
The configuration is completely equivalent to the elaborated specification below:

```{yaml}
variables:
  HW00002:
    name: "subject_height"
    type: "numeric"
    bounds:
      min: 0.75
      max: 2.25
  HW00003:
    name: "subject_likes_cake"
    type: "categorical"
    levels:
      "1":
        name: "no"
        alternate_patterns:
          - "0"
      "2":
        name: "yes"
        alternate_patterns:
          - "1"
    "na-values":
      - "none"
      - "unknown"
      - "not available"
```

Though the above examples are used in the dataset configuration `variables` section,
they are also valid to use in the `derived` variables section.

```{yaml}
derived:
  cake_superfan:
    name: "subject really likes cake a lot"
    shared_model: "yesno"
    code: "HW00003"
```

## Overriding Default Shared Model Settings

Shared models are designed to allow individual variables that invoke the shared
model to override their default settings with variable-specific values. Note that
while this is supported, this isn't necessarily desirable if the goal is
harmonization.

### Use a model, but specify a custom set of `na-values`

Consider the following shared model and variable specification:

```{yaml}
models:
  testmodel:
    type: "string"
    "na-values":
      - "general_value"
```

```{yaml}
variables:
  HW00001:
    name: "testvar"
    shared_model: "testmodel"
    "na-values":
      - "specific_value"
```

The resolution of this shared model would be:

```{yaml}
variables:
  HW00001:
    name: "testvar"
    type: "string"
    "na-values":
      - "specific_value"
```

### Use a model, but partially modify the categorical levels

Consider the following example:

```{yaml}
models:
  testmodel:
    type: "categorical"
    levels:
      "1":
        name: "lvl1"
        alternate_patterns:
          - "pat1"
      "2":
        name: "lvl2"
        alternate_patterns:
          - "pat2"
```

```{yaml}
variables:
  HW00001:
    name: "testvar"
    shared_model: "testmodel"
    levels:
      "2":
        alternate_patterns:
          - "patoverride"
      "3":
        name: "lvl3"
        alternate_patterns:
          - "pat3"
```

This is a rather toxic example, as the resulting variable would
have a hybrid categorical model that, when encoded as a factor, wouldn't
be particularly compatible with other variables that use this model.
However, in the interest of curiosity, the final variable configuration
would be resolved as follows:

```{yaml}
variables:
  HW00001:
    name: "testvar"
    type: "categorical"
    levels:
      "1":
        name: "lvl1"
        alternate_patterns:
          - "pat1"
      "2":
        name: "lvl2"
        alternate_patterns:
          - "patoverride"
      "3":
        name: "lvl3"
        alternate_patterns:
          - "pat3"
```


## Shared Models for SurveyCTO Data

SurveyCTO form definitions by default come with a `choices` tab that specifies
categorical models for use in the survey. The utility function `process.phenotypes::parse.surveycto`
will automatically generate a shared model file based on the `choices` tab of a form.
If desired, additional shared model data can be manually entered into the shared model
configuration file after generation. Please see
[Configuration from a SurveyCTO Form Definition](configuration-from-SurveyCTO-form.html)
for more information.
