---
title: "Validation of Dataset Configuration Files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Validation of Dataset Configuration Files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(process.phenotypes)
```

To facilitate the process of creating configuration files for a phenotype dataset,
`process.phenotypes` contains a utility function and script for config validation. This step
is completely optional, but can help catch some of the most common discrepancies
in manually configuration.

## Validator overview

The yaml validator will not catch all possible discrepancies in configuration
files, but it will flag the following:

- `tag`
- missing required properties from globals
- missing required properties from each variable definition in `variables` and `derived`
- malformed types
  - accepted types are: `string`, `binary`, `categorical`, `categorical_to_numeric`, `ordinal`,
    `numeric`, `blood pressure` or `blood_pressure` or `bloodpressure` or `bp`, `date`
  - any capitalization of any type name is accepted
- invalid variable configuration with either `type` or `shared_model`

The yaml validator will *not* catch instances of duplicate keys in yaml maps (e.g. names of entries
in `variables`, `derived`, or `levels`). These will instead flag as an error when executing
`process.phenotypes::create.phenotype.report`, and can be corrected at that time.

## Install necessary packages

The suggested packages `jsonvalidate` and `rjson` are required for this utility.
They are available in conda as `r-jsonvalidate` and `r-rjson`.

```{r load.input.libraries}
library(jsonvalidate)
library(rjson)
library(yaml)
```

## Validator usage

Assume that you have a dataset configuration yaml named `dataset.yaml`:

```{yaml}
tag: "HW"
globals:
  "min_age_for_inclusion": 16
  "max_invalid_datatypes_per_subject": 10
  "consent_inclusion_file": ~
  "consent_exclusion_file": ~
variables:
  "HW00001":
    "name": "var1"
    "type": "string"
    "canonical_name": "variable 1"
    "subject_id": yes
  "HW00002":
    "name": "var2"
    "type": "numeric"
    "canonical_name": "variable 2"
    "subject_age": yes
  "HW00003":
    "name": "var3"
    "shared_model": "yesno"
    "canonical_name": "variable 3"
derived:
  "harmonized_name":
    "name": "aliased var3"
    "shared_model": "yesno"
    "code": "HW00003"
```

and a shared models configuration yaml named `shared-models.yaml`:

```{yaml}
models:
  "yesno":
    "type": "categorical"
    "levels":
      "1":
        "name": "no"
      "2":
        "name": "yes"
```

Then schema validation can be run as follows:

```{R run.good.validation, eval=TRUE, echo=TRUE}
process.phenotypes::config.validation(
  "dataset-valid.yaml", "shared-models-valid.yaml",
  system.file("validator/schema.datasets.yaml", package = "process.phenotypes"),
  system.file("validator/schema.shared-models.yaml", package = "process.phenotypes")
)
```

### Interpretation of validation errors

Validator errors tend to be rather verbose, and require some interpretation.
If either of the above notifications is missing, then something about one of the configuration
files does not match expectation. The informative message will be somewhere near the top of the
error, likely accompanied by an assortment of tangentially related errors the validator
encountered when attempting to resolve the configuration file.

Consider the following malformed version of the above dataset yaml:

```{yaml}
tag: "HW"
globals:
  "min_age_for_inclusion": 16
  "max_invalid_datatypes_per_subject": 10
  "consent_inclusion_file": ~
  "consent_exclusion_file": ~
variables:
  "HW00001":
    "name": "var1"
    "type": "string"
    "canonical_name": "variable 1"
    "subject_id": yes
  "HW00002":
    "name": "var2"
    "type": "numeric"
    "canonical_name": "variable 2"
    "subject_age": yes
  "HW00003":
    "name": "var3"
    "shared-model": "yesno"
    "canonical_name": "variable 3"
derived:
  "harmonized_name":
    "name": "aliased var3"
    "shared_model": "yesno"
    "code": "HW00003"
```

The tag `shared-model` is incorrect; the schema expects `shared_model`. If we run
this configuration file through the validator:

```{R run.bad.validation, eval=TRUE, echo=TRUE, error=TRUE}
process.phenotypes::config.validation(
  "dataset-invalid.yaml", "shared-models-valid.yaml",
  system.file("validator/schema.datasets.yaml", package = "process.phenotypes"),
  system.file("validator/schema.shared-models.yaml", package = "process.phenotypes")
)
```

In this case, the error `... must have required property 'shared_model'` indicates
that the configuration block was unable to match the `shared_model` block specification,
since the critical tag was malformed.
