---
title: "Phenotype Quality Control Report"
output:
  html_document:
    code_folding: hide
    md_extensions: "-fancy_lists -emoji"
  highlight: tango
  number_sections: no
  theme: default
  toc: yes
  toc_depth: 3
  toc_float:
    collapsed: no
    smooth_scroll: yes
params:
  dataset.name: ""
  variable.summary: NULL
  phenotype.data: ""
  unique.variable.value.inclusion.proportion: 0.333
  subjects.wrong.type: c()
  variables.wrong.type: c()
  nas.by.subject: c()
  subjects.failing.deps: c()
  subj.invalid.type.max: 100
---

```{r link.variables, eval=TRUE, echo=FALSE}
#### Link input parameters to local variables
dataset.name <- params$dataset.name
variable.summary <- params$variable.summary
phenotype.data <- params$phenotype.data
unique.var.value.inc.prop <- params$unique.variable.value.inclusion.proportion
subjects.wrong.type <- params$subjects.wrong.type
variables.wrong.type <- params$variables.wrong.type
nas.by.subject <- params$nas.by.subject
subjects.failing.deps <- params$subjects.failing.deps
subj.invalid.type.max <- params$subj.invalid.type.max
```

```{r load.packages, eval=TRUE, echo=FALSE}
#### Load required R packages
suppressPackageStartupMessages(library(ggplot2, quietly = TRUE))
suppressPackageStartupMessages(library(RColorBrewer, quietly = TRUE))
suppressPackageStartupMessages(library(knitr, quietly = TRUE))
suppressPackageStartupMessages(library(kableExtra, quietly = TRUE))
```

```{r get.bins, eval=TRUE, echo=FALSE}
#### Make sure the number of bins in a histogram isn't too close to the number of unique values!
get.bins <- function(vec) {
  n.unique.values <- length(unique(vec))
  if (n.unique.values > 50) {
    50
  } else {
    n.unique.values
  }
}
```

***
<br>

```{r configure.ggplot2.theme, eval=TRUE, echo=FALSE}
#### Configure default themes for plotting/tables
my.theme <- theme_light() + theme(
  plot.title = element_text(size = 16, hjust = 0.5),
  axis.title = element_text(size = 14),
  axis.text = element_text(size = 12),
  strip.background = element_blank(),
  strip.text = element_text(size = 14, colour = "black"),
  legend.title = element_text(size = 14),
  legend.text = element_text(size = 13)
)
```

This report and the accompanying cleaned data is generated by the code available 
at <a href="https://gitlab.com/data-analysis5/process.phenotypes">this</a> repository, 
run using the configuration outlined in `r dataset.yaml`.
<br><br>
The report is a summary of the phenotype data and includes subject-level summary metrics, 
variable-level summary metrics, a linker between original variable names and standardized encoded names,
and per-variable summaries.  The per-variable summaries vary based on data type.  Numeric types will
display a histogram and min/max/mean/median/deciles.  Categorical and ordinal types display counts of each 
factor level.  All types other than string include a summary of data entries that are not consistent with 
the indicated data type.  Dependencies, bounds, data types, etc. are defined in the accompanying configuration
file.  
<br>
Please note that with few exceptions, the original variable names and descriptions in this report are read
verbatim from the input data, and have not been edited, to maintain harmonization with the original questionnaire
definitions.  
<br>
`r paste(variable.summary$total.initial.sample.size, ifelse(variable.summary$total.initial.sample.size == 1, " subject was", " subjects were"), " originally present in this dataset.", ifelse(!is.null(variable.summary$globals$consent_inclusion_file), paste(" ", variable.summary$subjects.consent.yes, ifelse(variable.summary$subjects.consent.yes == 1, " subject was", " subjects were"), " included with verified consent.", sep = ""), ""), ifelse(!is.null(variable.summary$globals$consent_exclusion_file), paste(" ", variable.summary$subjects.consent.no, ifelse(variable.summary$subjects.consent.no == 1, " subject was", " subjects were"), " excluded for lacking consent.", sep = ""), ""), ifelse(!is.null(variable.summary$globals$consent_inclusion_file) & !is.null(variable.summary$globals$consent_exclusion_file), paste(" ", length(variable.summary$subjects.ambiguous.consent), ifelse(length(variable.summary$subjects.ambiguous.consent) == 1, " subject was", " subjects were"), " additionally excluded for not being explicitly listed as either having or lacking consent.", sep = ""), ""), " ", variable.summary$subjects.excluded.for.age, ifelse(variable.summary$subjects.excluded.for.age == 1, " subject was ", " subjects were "), "removed for being below the minimum permissible age of ", variable.summary$globals$min_age_for_inclusion, ".", sep = "")`

`r paste(variable.summary$na.subject.id.count, ifelse(variable.summary$na.subject.id.count == 1, " subject was ", " subjects were "), "removed for lacking a subject ID.", sep = "")`

`r if (length(variable.summary$subjects.ambiguous.consent) > 0) {paste("***\n<br>")}`

```{r report.consent.ambiguity, eval=length(variable.summary$subjects.ambiguous.consent) > 0, echo=FALSE, results="asis"}
ambig.consent <- data.frame(x = sort(variable.summary$subjects.ambiguous.consent))
colnames(ambig.consent) <- c("Subject ID")
knitr::kable(ambig.consent, caption = "Subjects Absent from Consent Tracking") %>%
  kableExtra::kable_styling("condensed", position = "left", full_width = FALSE)
```

***
<br>

```{r report.invalid.type.removal, eval=TRUE, echo=FALSE, results="asis"}
n.invalid.types.removed <- 0
if (!is.null(subj.invalid.type.max)) {
  n.invalid.types.removed <- length(which(subjects.wrong.type >= subj.invalid.type.max))
  cat("\n",
    n.invalid.types.removed,
    ifelse(n.invalid.types.removed == 1, "subject", "subjects"),
    "will be removed from cleaned output data for containing more than",
    subj.invalid.type.max,
    "datapoints not matching the corresponding variable type.\n",
    sep = " "
  )
}
```
<br>
`r as.integer(nrow(phenotype.data) - n.invalid.types.removed)` subjects remain in the cleaned dataset.
Based on configured dependency relationships between variables,
`r variable.summary$actual.nas.from.deps` of `r variable.summary$possible.nas.from.deps`,
or `r ifelse(variable.summary$possible.nas.from.deps == 0, 0, round(variable.summary$actual.nas.from.deps / variable.summary$possible.nas.from.deps, 2))`%,
of possible datapoint exclusions `r ifelse(variable.summary$actual.nas.from.deps == 1, "was", "were")` set to NA
due to failed dependencies.

***
<br>

### Subject level metrics

```{r report.subject.specific.na, eval=TRUE, echo=FALSE, results="asis"}
## probably want a histogram of this one
plot.data <- data.frame(x = nas.by.subject)
my.hist <- ggplot(aes(x = x, y = ..count.. / sum(..count..)), data = plot.data)
nbins <- get.bins(nas.by.subject)
my.hist <- my.hist + my.theme + geom_histogram(bins = nbins, binwidth = (diff(range(nas.by.subject)) + 1) / nbins)
my.hist <- my.hist + xlab("Number of NA Values by Subject") + ylab("Proportion of Subjects")
cat("\n\n#### Distribution of NA Counts by Subject\n\n")
print(my.hist)

knitr::kable(process.phenotypes:::get.top.ten(TRUE, nas.by.subject, "Subjects"),
  caption = "Ten highest NA-count subjects"
) %>%
  kableExtra::kable_styling("condensed", position = "left", full_width = FALSE)
```

```{r report.subject.type.failures, eval=length(subjects.wrong.type) > 0, echo=FALSE, results="asis"}
## probably also want a histogram of this one
plot.data <- data.frame(x = subjects.wrong.type)
my.hist <- ggplot(aes(x = x, y = ..count.. / nrow(phenotype.data)), data = plot.data)
nbins <- get.bins(subjects.wrong.type)
my.hist <- my.hist + my.theme + geom_histogram(bins = nbins, binwidth = (diff(range(subjects.wrong.type)) + 1) / nbins)
my.hist <- my.hist + xlab("Number of Invalid Type Datapoints by Subject") + ylab("Proportion of Subjects")
my.hist <- my.hist + geom_vline(xintercept = subj.invalid.type.max, col = "red", lty = 2)
cat("\n\n#### Distribution of Invalid Type Datapoints by Subject\n\n")
print(my.hist)

knitr::kable(process.phenotypes:::get.top.ten(TRUE, subjects.wrong.type, "Subjects"),
  caption = "Ten highest type conversion failures"
) %>%
  kableExtra::kable_styling("condensed", position = "left", full_width = FALSE)
```

***
<br>

```{r report.subject.dependency.failures, eval=length(which(subjects.failing.deps > 0)) > 0, echo=FALSE, results="asis"}
## histogram time
plot.data <- data.frame(x = subjects.failing.deps)
my.hist <- ggplot(aes(x = x, y = ..count.. / nrow(phenotype.data)), data = plot.data)
nbins <- get.bins(subjects.failing.deps)
my.hist <- my.hist + my.theme + geom_histogram(
  bins = nbins,
  binwidth = (diff(range(subjects.failing.deps)) + 1) / nbins
)
my.hist <- my.hist + xlab("Number of Variable Dependency Failures by Subject") + ylab("Proportion of Subjects")
cat("\n\n#### Distribution of Dependency Failures by Subject\n\n")
print(my.hist)
```

***
<br>

`r if (length(which(variables.wrong.type > 0)) > 0) {"### Variable level metrics"}`

```{r report.variable.type.failures, eval=length(which(variables.wrong.type > 0)) > 0, echo=FALSE, results="asis"}
## for now, histogram this too
plot.data <- data.frame(x = variables.wrong.type)
my.hist <- ggplot(aes(x = x, y = ..count.. / sum(..count..)), data = plot.data)
nbins <- get.bins(variables.wrong.type)
my.hist <- my.hist + my.theme + geom_histogram(bins = nbins, binwidth = (diff(range(variables.wrong.type)) + 1) / nbins)
my.hist <- my.hist + xlab("Number of Invalid Type Datapoints by Variable") + ylab("Proportion of Variables")
cat("\n\n#### Distribution of Invalid Type Datapoints by Variable\n\n")
print(my.hist)

knitr::kable(process.phenotypes:::get.top.ten(TRUE, variables.wrong.type, "Variables"),
  caption = "Ten highest type conversion failures"
) %>%
  kableExtra::kable_styling("condensed", position = "left", full_width = FALSE)
```

### Report linking between input and standardized variable names

```{r report.variable.mappings, eval=TRUE, echo=FALSE, results="asis"}
variable.mapping.df <- data.frame(
  names(variable.summary$variables),
  sapply(variable.summary$variables, function(i) {
    i$original.name
  }),
  sapply(variable.summary$variables, function(i) {
    if (!is.null(i$params$canonical_name)) {
      i$params$canonical_name
    } else {
      ""
    }
  })
)
rownames(variable.mapping.df) <- NULL
colnames(variable.mapping.df) <- c("Standardized", "Original", "Description")
knitr::kable(variable.mapping.df, caption = "Phenotype Variable Names") %>%
  kableExtra::kable_styling("condensed", position = "left", full_width = FALSE)
```

***
<br>

### Report variable distribution summaries

In this section, each variable in turn is summarized according
to the variable's configured type.

***
<br>

```{r report.value.summary, eval=TRUE, echo=FALSE, results="asis"}
for (name in names(variable.summary$variables)) {
  my.summ <- variable.summary$variables[[name]]$summary
  variable.pretty.name <- variable.summary$variables[[name]]$original.name
  suppress.reporting <- ifelse(!is.null(variable.summary$variables[[name]]$params$suppress_reporting),
    variable.summary$variables[[name]]$params$suppress_reporting,
    FALSE
  )
  cat(
    "\n\n### ", variable.pretty.name, "\n\n"
  )
  report.name.and.code(variable.summary$variables[[name]])

  ## report any indication of residual Excel junk in the variable
  report.excel.problems(
    variable.summary$variables[[name]],
    suppress.reporting
  )


  ## if the source data type is numeric
  report.numeric.summary(
    phenotype.data[, name],
    variable.summary$variables[[name]],
    name,
    variable.pretty.name,
    suppress.reporting
  )

  ## if the source data has a linked_date variable, indicating it is an age
  ## that in theory should be derived from a date
  report.linked.date(
    phenotype.data[, name],
    phenotype.data,
    variable.summary$variables[[name]],
    name,
    suppress.reporting
  )
  ## if the source data has a computed_bmi variable, indicating it is BMI
  ## that in theory should be derived from height and weight
  report.bmi.comparison(
    phenotype.data,
    variable.summary$variables[[name]],
    name,
    suppress.reporting
  )
  ## if the source data has a computed_bp_ratio variable, plot the ratio
  ## between the systolic and diastolic measurements as indicated
  report.bp.ratio(
    phenotype.data,
    variable.summary$variables[[name]],
    name,
    suppress.reporting
  )

  ## original exclusion criterion here is hitting everything for small datasets,
  ## even when the report is not a unique value report
  report.content.summary(
    phenotype.data,
    my.summ,
    unique.var.value.inc.prop,
    name,
    variable.pretty.name,
    suppress.reporting
  )
  ## blood pressure entries not matching expected format are generally reported,
  ## as deviations from accepted formats are very common
  report.noncompliant.bp(
    variable.summary$variables[[name]],
    name,
    suppress.reporting
  )

  report.noncompliant.numerics(
    variable.summary$variables[[name]],
    name,
    suppress.reporting
  )
  ## factors
  report.factor.summary(
    variable.summary$variables[[name]],
    name,
    suppress.reporting
  )

  report.noncompliant.dates(
    variable.summary$variables[[name]],
    name,
    suppress.reporting
  )
  report.unicode.entries(
    variable.summary$variables[[name]],
    suppress.reporting
  )

  report.dependencies(
    variable.summary,
    name,
    suppress.reporting
  )

  cat("\n***\n<br>\n")
}
```

***
<br>

#### Session Information

<br>

The following summarizes the loaded R configuration for the run that created this report.

```{r session.info, eval=TRUE, echo=TRUE}
sessionInfo()
```
